---
title: "Introduction to MECfda Package"
author: 
  - name: "Heyang Ji"
output: rmarkdown::html_vignette
abstract: 
  "Package MECfda provide functions that sovle a variaty of scalar-on-function linear regression models
  and estimation methods that correct the bias due to measurement error in such models. 
  "
vignette: >
  %\VignetteIndexEntry{Introduction to MECfda Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<!-- output: pdf_document -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Scalar-on-function linear regression is an important problem in functional data analysis.
Our aim is to develop an R package that can solve all kinds of scalar-on-function linear regression models and 
correct the bias due to measurement error in scalar-on-function linear regression models. 

# Scalar-on-function Linear Regression

The generic form of scalar-on-function linear regression model is
$$\delta(F_{Y_i|X_i,Z_i}) =  \sum_{l=1}^{L} \int_{\Omega_l} \beta_l(t) X_{li}(t) dt + \sum_{j=1}^{J} \gamma_j Z_{ji} + \gamma_0$$
where 
$Y_i$ represents the scalar-valued response variable, 
$X_li(t)$ represents the function-valued covariates ($l=1,\dots,L$), 
$Z_ji$ represents the scalar valued covariates ($j=1,\dots,J$),
$F_{Y_i|X_i,Z_i}$ represents the CDF of the $Y_i$ given $X_i$ and $Z_i$, 
$\delta(\cdot)$ is a functional. 
In ordinary scalar-on-function generalized linear (mixed) models,
$$\delta(F_{Y_i|X_i,Z_i}) = g\left\{\int_\mathbb{R}ydF_{Y_i|X_i,Z_i}(y)\right\} = g(E[Y_i|X_i,Z_i]),$$ 
where $g(\cdot)$ is a link function. 
In scalar-on-function linear quantile regression models,
$$\delta(F_{Y_i|X_i,Z_i}) = F_{Y_i|X_i,Z_i}^{-1}(\tau),$$ 
where $\tau\in(0,1)$. 

## Basis Expansion



```{r setup}
library(MECfda)
```
